A full approximation storage multilevel method \\ for nonlinear variational inequalities

### 2023-07-04 paper/fascd.tex
Edits through Section 7

When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.  Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.

@@

In the material which follows, use of higher-order Lagrange elements, i.e.~$P_k$ or $Q_k$ with $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to inequality constraints, which we avoid.  Namely, our FE construction of constraint sets (Section \ref{sec:cdmultilevel}) will require \emph{nodal monotonicity},
\begin{equation}
\bw \ge \bz \quad \implies \quad w \ge z \label{eq:nodalmonotonicity}
\end{equation}
for all $w,z \in \mathcal{V}^j$, where $\bw,\bz \in \RR^{m_j}$ denote vectors of nodal values.  However, for $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal values one or zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.  Evidently \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of such indefinite-sign hat functions.
\pef{I'm not sure this constitutes a variational crime? We enforce the RHS because it's tractable, and this is only equivalent to what we actually want (the LHS) for P1 and Q1. I don't see where we're departing from the variational framework, however.}

@@

\begin{pseudofloat}[ht]
\begin{pseudo}
\pr{fascd-vcycle}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J)\text{:} \\+
    $\underline{\chi}^J, \, \overline{\chi}^J = \underline{\gamma}^J - w^J, \, \overline{\gamma}^J - w^J {\large \strut}$ \label{line:vcyclegenchifinest} \\
    for $j=J$ downto $j=1$ \\+
      $\underline{\chi}^{j-1}, \, \overline{\chi}^{j-1} = \maxR \underline{\chi}^j, \, \minR \overline{\chi}^j {\large \strut}$ \label{line:vcyclegenchi} \\
      $\underline{\phi}^j, \, \overline{\phi}^j = \underline{\chi}^j - P\underline{\chi}^{j-1}, \, \overline{\chi}^j - P\overline{\chi}^{j-1} {\large \strut}$ \label{line:vcyclegenphi} \\
      $y^j = 0$ \\
      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,\overline{\phi}^j,w^j;y^j)$  \ct{smoothing of \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
      $w^{j-1} = \iR(w^j + y^j)$ \label{line:vcyclerestrictsolution} \\
      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \label{line:vcyclerestrictell} \\-
    $z^0 = 0$ \\
    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,w^0;z^0)$ \hspace{1.0cm} \ct{solving of \eqref{eq:fe:upvi} in $\mathcal{U}^0$} \\
    for $j=1$ to $j=J$ \\+
      $z^j = y^{j} + P z^{j-1}$ \\
      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,w^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
    return $w^J+z^J$
\end{pseudo}
\caption{The FASCD V-cycle is an iteration for solving FE VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
\label{alg:fascd}
\end{pseudofloat}

Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle.\footnote{We could write ``$V(\id{down},\id{up})$'', but we prefer to emphasize the number of subsets used on each level rather than the number of smoother sweeps in a given set. \pef{I don't understand this.}}  If \id{up} $=0$ then it becomes a $V(1,0)$ cycle and if \id{down} $=0$ then it is a $V(0,1)$ cycle; in the former case it generalizes Algorithm 4.7 in \cite{GraeserKornhuber2009} by allowing nonlinear operators $f$, any smoother, and applying FAS-type corrections.  In unilateral cases it is precise to say that one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over incomplete CD \eqref{eq:fe:upwardsuminclusion}; in this sense $V(1,1)=V(1,0)+V(0,1)$.

Regarding the floating-point storage required, let us assuming all box constraints are nontrivial.  Then seven vectors must be allocated on each level: $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $w^j$, $\ell^j$, $y^j$.  Note $z^j$ can use the same storage as $y^j$.  On the finest level one must also store $\underline{\gamma}^J$ and $\overline{\gamma}^J$, while on the coarsest level there are only five vectors because $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.  Therefore, using $m_j=2^d m_{j-1}$, plus a geometric series argument, the total storage, excluding smoother and coarse-solver implementations, is
\begin{equation}
9 m_J + 7 m_{J-1} + \dots + 7 m_1 + 5 m_0 \le \left(9 + \frac{7}{2^d - 1}\right) m_J.
\end{equation}
This gives storage at most $16m_J$, $12m_J$, and $10m_J$ in dimensions $d=1,2,3$, respectively.  For comparison, a single-level method needs $4 m_J$ storage (i.e.~$\underline{\gamma}^J,\overline{\gamma}^J,\ell^J,w^J$).  If some of the obstacles are trivial, for instance in the unilateral case $\overline{\gamma}^J=+\infty$, storage can be correspondingly reduced at all levels.

@@

Because the problem is inequality-constrained, repeated application of Algorithm \ref{alg:fascd} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.  To construct an appropriate convergence criteria for the discretization, note \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, a strong form.  Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding hat function.  Let $\mathcal{X}^J=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}^J$.  If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let $I_D$ be the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.  Finally, write $a\perp b$ when $a,b \in \RR^n$ are \emph{complementary}, that is, when $a_i \ge 0$ and $b_i \ge 0$ and $a_i b_i = 0$ for $i=0,\dots,n-1$.  Then problem \eqref{eq:fe:vi}, which would be expressed VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap \mathcal{X}^J\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP for the solution $u^J \in \mathcal{K}^J$, but with additional equality constraints:
\begin{align}
\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
I_D u^J &= g_D^J\big|_{\partial_D\Omega} \notag
\end{align}
for some Lagrange multipliers $\lambda \in \RR^{m_D}$.

@@

@@

Suppose an iteration starts with $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.  Let $\id{atol}\ge 0$, $\id{rtol}\ge 0$ be given tolerances.  We propose
\begin{equation}
\|\rSS(w^{J,k})\|_2 < \id{atol} \qquad \text{or} \qquad \frac{\|\rSS(w^{J,k})\|_2}{\|\rSS(w^{J,0})\|_2} < \id{rtol} \label{eq:stoppingcriterion}
\end{equation}
as the stopping criterion.  A step tolerance, $\|w^{J,k} - w^{J,k-1}\|_2 < \id{stol}$ or similar, could also be used, but for the results in Section \ref{sec:results} we did not find it necessary.

Iteration of V-cycle Algorithm \ref{alg:fascd}, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, will be shown (Section \ref{sec:results}) to be an effective solver.  However, an F-cycle (Figure \ref{fig:fcycle}), also called a \emph{full multigrid cycle}, is even better.  The well-known principle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate.  Then an admissible initial iterate on the next-finer mesh is generated via prolongation and truncation, and so on.  During this ``ramp'' up one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh.  From the finest level one does full-depth V-cycles to convergence.

\begin{figure}[h]
\begin{center}
\input{tikz/fcycle.tex}
\end{center}
\caption{F-cycle Algorithm \ref{alg:fascd-fcycle} initially restricts the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray edges).  Then it prolongs and truncates (doubled edges) to create an admissible initial iterate on the next-finer level.  Solution on each level is by V-cycles; \id{rampv} $=1$ here.  See Figure \ref{fig:fascdvcycle} for the other symbols.}
\label{fig:fcycle}
\end{figure}

This F-cycle approach is in Algorithm \ref{alg:fascd-fcycle}.  Observe that while injection of the original fine-level constraints $\underline{\gamma}^J, \overline{\gamma}^J$ is \emph{not} used for V-cycles, wherein the LDCs use an existing admissible iterate (Section \ref{sec:cdmultilevel}), when combined with truncation into $\mathcal{K}^j=\{v \in \mathcal{V}^j \,:\, \underline{\gamma}^j \le v \le \overline{\gamma}^j, \, v|_{\partial\Omega} = g_D\}$ injection is appropriate for the ramp in an F-cycle; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo}.  One may generate the coarsest-level initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation down to coarsest level; compare line \ref{line:fcyclecoarsestinitial}.

\begin{pseudofloat}[h]
\begin{pseudo}
\pr{fascd-fcycle}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J)\text{:} \\+
    for $j=J$ downto $j=1$ \\+
        $\ell^{j-1} = R(\ell^j)$ \\
        $\underline{\gamma}^{j-1}, \, \overline{\gamma}^{j-1} = \iR \underline{\gamma}^{j}, \, \iR \overline{\gamma}^{j}$ {\large\strut} \\-
    choose $w^0 \in \mathcal{K}^0$ \label{line:fcyclecoarsestinitial} \\
    $\text{\pr{solve}}(\ell^0,\underline{\gamma}^0,\overline{\gamma}^0;w^0)$ \\
    for $j=1$ to $j=J-1$ \\+
        $w^j = \max\left\{\underline{\gamma}^{j},\min\{\overline{\gamma}^{j}, Pw^{j-1}\}\right\} \in \mathcal{K}^j$ \label{line:fcycleprolongtruncone} \\
        $\pr{fascd-vcycle}^{\text{\id{rampv}}}(j,\ell^j,\underline{\gamma}^j,\overline{\gamma}^j;w^j)$ \\-
    $w^J = \max\left\{\underline{\gamma}^{J},\min\{\overline{\gamma}^{J}, Pw^{J-1}\}\right\} \in \mathcal{K}^J$ \label{line:fcycleprolongtrunctwo} \\
    while not \eqref{eq:stoppingcriterion} \\+
        \pr{fascd-vcycle}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J) \\-
    return $w^J$
\end{pseudo}
\caption{The FASCD F-cycle for solving FE VI problem \eqref{eq:fe:vi}.}
\label{alg:fascd-fcycle}
\end{pseudofloat}

@@

A large space of smoothers could be explored,\footnote{See comments in Section \ref{sec:discussion}.} but Section \ref{sec:results} only demonstrates one powerful, $O(m_j)$ work per iteration, smoother choice.  We apply a fixed number of iterations of the active-set (reduced-space) Newton method from \cite{BensonMunson2006}, the \texttt{-snes\_type vinewtonrsls} solver in PETSc \cite{Balayetal2023}, denoted as \emph{RS Newton}.  We use it with a fixed number of preconditioned Krylov iterations to solve the step equations.  For the symmetric Laplacian in Example \ref{ex:results:classical} we apply conjugate gradients (CG) with incomplete Cholesky (IC) preconditioning, but for other operators we apply the generalized minimum residual (GMRES) method with incomplete LU (ILU) preconditioning.  The coarse solver is also RS Newton, but iterated to convergence and with LU decomposition solution of the step equations.

@@

\medskip\noindent
Ed: \cite{GraeserKornhuber2009} only considers the Laplacian lower obstacle case, and they write everything in terms of nonlinear Gauss-Seidel (i.e. optimize one hat function at a time).  They first lay out the \cite{Tai2003} CD strategy as Alg. 4.7, only for V(1,0) cycles.  Then they observe that the convergence rates from \cite{Tai2003} are slower (worse) than if one does "monotone multigrid".  This seems to mean to them that one modifies the coarse hat functions according to whether they overlap the current fine-grid estimate of the coincidence set, "truncated nodal basis functions".  (See the start of \cite{GraeserKornhuber2009} section 5.2.)  I think that making these modifications in Firedrake, so we could try monotone multigrid V-cycles, would be a nightmare.  FASCD does not attempt the monotone multigrid idea at all.

>>>>

When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over basis function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.  Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.

@@

In the material which follows, use of higher-order Lagrange elements, i.e.~$P_k$ or $Q_k$ with $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to inequality constraints, which we avoid.  Namely, our FE construction of constraint sets (Section \ref{sec:cdmultilevel}) will require \emph{nodal monotonicity},
\begin{equation}
\bw \ge \bz \quad \implies \quad w \ge z \label{eq:nodalmonotonicity}
\end{equation}
for all $w,z \in \mathcal{V}^j$, where $\bw,\bz \in \RR^{m_j}$ denote vectors of nodal values.  However, for $k\ge 2$ there are basis functions in $P_k$ and $Q_k$, i.e.~functions with nodal values one or zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.  Evidently \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of such indefinite-sign basis functions.
\pef{I'm not sure this constitutes a variational crime? We enforce the RHS because it's tractable, and this is only equivalent to what we actually want (the LHS) for P1 and Q1. I don't see where we're departing from the variational framework, however.}

@@

\begin{pseudofloat}[ht]
\begin{pseudo}
\pr{fascd-vcycle}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J)\text{:} \\+
    $\underline{\chi}^J, \, \overline{\chi}^J = \underline{\gamma}^J - w^J, \, \overline{\gamma}^J - w^J {\large \strut}$ \label{line:vcyclegenchifinest} \\
    for $j=J$ downto $j=1$ \\+
      $\underline{\chi}^{j-1}, \, \overline{\chi}^{j-1} = \maxR \underline{\chi}^j, \, \minR \overline{\chi}^j {\large \strut}$ \label{line:vcyclegenchi} \\
      $\underline{\phi}^j, \, \overline{\phi}^j = \underline{\chi}^j - P\underline{\chi}^{j-1}, \, \overline{\chi}^j - P\overline{\chi}^{j-1} {\large \strut}$ \label{line:vcyclegenphi} \\
      $y^j = 0$ \\
      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,\overline{\phi}^j,w^j;y^j)$  \ct{smoothing of \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
      $w^{j-1} = \iR(w^j + y^j)$ \label{line:vcyclerestrictsolution} \\
      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \label{line:vcyclerestrictell} \\-
    $z^0 = 0$ \\
    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,w^0;z^0)$ \hspace{1.0cm} \ct{solving of \eqref{eq:fe:upvi} in $\mathcal{U}^0$} \\
    for $j=1$ to $j=J$ \\+
      $z^j = y^{j} + P z^{j-1}$ \\
      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,w^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
    return $w^J+z^J$
\end{pseudo}
\caption{The FASCD V-cycle is an iteration for solving VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
\label{alg:fascd}
\end{pseudofloat}

Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle.\footnote{We could write ``$V(\id{down},\id{up})$'', but we prefer to emphasize the number of subsets used on each level rather than the number of smoother sweeps in a given set. \pef{I don't understand what the number of subsets used on each level means.}}  If \id{up} $=0$ then it becomes a $V(1,0)$ cycle and if \id{down} $=0$ then it is a $V(0,1)$ cycle; in the former case it generalizes Algorithm 4.7 in \cite{GraeserKornhuber2009} by allowing nonlinear operators $f$, any smoother, and applying FAS-type corrections.  In unilateral cases, one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over incomplete CD \eqref{eq:fe:upwardsuminclusion}; in this sense $V(1,1)=V(1,0)+V(0,1)$.

Regarding the storage required, let us assuming all box constraints are nontrivial. In this case, seven vectors must be allocated on each level: $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $w^j$, $\ell^j$, $y^j$.  Note $z^j$ can use the same storage as $y^j$.  On the finest level one must also store $\underline{\gamma}^J$ and $\overline{\gamma}^J$, while on the coarsest level there are only five vectors because $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.  Therefore, using $m_j=2^d m_{j-1}$, plus a geometric series argument, the total storage, excluding smoother and coarse-solver implementations, is
\begin{equation}
9 m_J + 7 m_{J-1} + \dots + 7 m_1 + 5 m_0 \le \left(9 + \frac{7}{2^d - 1}\right) m_J.
\end{equation}
This gives storage at most $16m_J$, $12m_J$, and $10m_J$ in dimensions $d=1,2,3$, respectively.  For comparison, a single-level method needs $4 m_J$ storage (i.e.~$\underline{\gamma}^J,\overline{\gamma}^J,\ell^J,w^J$).  If some of the obstacles are trivial, for instance in the unilateral case $\overline{\gamma}^J=+\infty$, storage can be correspondingly reduced at all levels.

@@

Because the problem is inequality-constrained, repeated application of Algorithm \ref{alg:fascd} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.  To construct appropriate convergence criteria for the discretization, note \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}.  Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding basis function.  Let $\mathcal{X}^J=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}^J$.  If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let $I_D$ be the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.  Finally, write $a\perp b$ when $a,b \in \RR^n$ are \emph{complementary}, that is, when $a_i \ge 0$ and $b_i \ge 0$ and $a_i b_i = 0$ for $i=0,\dots,n-1$.  Then problem \eqref{eq:fe:vi}, which would be expressed VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap \mathcal{X}^J\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP for the solution $u^J \in \mathcal{K}^J$, but with additional equality constraints:
\begin{align}
\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
I_D u^J &= g_D^J\big|_{\partial_D\Omega} \notag
\end{align}
for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
\pef{I wonder if this definition will confuse people, because they won't know how to plug in $\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ into your definition of $\perp$ ($\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ is not in $\RR^n$). Maybe it would be easier to just use the spelled-out version below, and not use this notation?}

@@

\pef{I think this is all fairly standard and we could abbreviate this section substantially.}

@@

Suppose an iteration starts with $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.  Let $\id{atol}\ge 0$, $\id{rtol}\ge 0$ be given tolerances.  We employ
\begin{equation}
\|\rSS(w^{J,k})\|_2 < \id{atol} \qquad \text{or} \qquad \frac{\|\rSS(w^{J,k})\|_2}{\|\rSS(w^{J,0})\|_2} < \id{rtol} \label{eq:stoppingcriterion}
\end{equation}
as the stopping criterion.  %A step tolerance, $\|w^{J,k} - w^{J,k-1}\|_2 < \id{stol}$ or similar, could also be used, but for the results in Section \ref{sec:results} we did not find it necessary.

In Section \ref{sec:results} we will show that iteration of the V-cycle Algorithm \ref{alg:fascd}, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, is an effective solver.
However, an F-cycle (Figure \ref{fig:fcycle}), also called a \emph{full multigrid cycle}, is even better.  The well-known principle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate.  Then an admissible initial iterate on the next-finer mesh is generated via prolongation and truncation\pef{Clarify truncation}, and so on.  During this ``ramp'' up one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh.  From the finest level one does full-depth V-cycles to convergence.

\begin{figure}[h]
\begin{center}
\input{tikz/fcycle.tex}
\end{center}
\caption{F-cycle Algorithm \ref{alg:fascd-fcycle} initially restricts the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray edges).  Then it prolongs and truncates (doubled edges) to create an admissible initial iterate on the next-finer level.  Solution on each level is by V-cycles. See Figure \ref{fig:fascdvcycle} for the other symbols.}
\label{fig:fcycle}
\end{figure}

This F-cycle approach is described in Algorithm \ref{alg:fascd-fcycle}.  Observe that while injection of the original fine-level constraints $\underline{\gamma}^J, \overline{\gamma}^J$ is \emph{not} used for V-cycles, wherein the LDCs use an existing admissible iterate (Section \ref{sec:cdmultilevel}), when combined with truncation into $\mathcal{K}^j=\{v \in \mathcal{V}^j \,:\, \underline{\gamma}^j \le v \le \overline{\gamma}^j, \, v|_{\partial\Omega} = g_D\}$, injection \emph{is} appropriate for the ramp in an F-cycle\pef{Explain why?}; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo}.  One may generate the coarsest-level initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation down to coarsest level; compare line \ref{line:fcyclecoarsestinitial}.

\begin{pseudofloat}[h]
\begin{pseudo}
\pr{fascd-fcycle}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J)\text{:} \\+
    for $j=J$ downto $j=1$ \\+
        $\ell^{j-1} = R(\ell^j)$ \\
        $\underline{\gamma}^{j-1}, \, \overline{\gamma}^{j-1} = \iR \underline{\gamma}^{j}, \, \iR \overline{\gamma}^{j}$ {\large\strut} \\-
    choose $w^0 \in \mathcal{K}^0$ \label{line:fcyclecoarsestinitial} \\
    $\text{\pr{solve}}(\ell^0,\underline{\gamma}^0,\overline{\gamma}^0;w^0)$ \\
    for $j=1$ to $j=J-1$ \\+
        $w^j = \max\left\{\underline{\gamma}^{j},\min\{\overline{\gamma}^{j}, Pw^{j-1}\}\right\} \in \mathcal{K}^j$ \label{line:fcycleprolongtruncone} \\
        $\pr{fascd-vcycle}^{\text{\id{rampv}}}(j,\ell^j,\underline{\gamma}^j,\overline{\gamma}^j;w^j)$ \\-
    $w^J = \max\left\{\underline{\gamma}^{J},\min\{\overline{\gamma}^{J}, Pw^{J-1}\}\right\} \in \mathcal{K}^J$ \label{line:fcycleprolongtrunctwo} \\
    while not \eqref{eq:stoppingcriterion} \\+
        \pr{fascd-vcycle}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J) \\-
    return $w^J$
\end{pseudo}
\caption{The FASCD F-cycle for solving VI problem \eqref{eq:fe:vi}.}
\label{alg:fascd-fcycle}
\end{pseudofloat}

@@

A large space of smoothers could be explored; in Section \ref{sec:results} we apply a fixed number of iterations of the active-set (reduced-space) Newton method from \cite{BensonMunson2006}, the \texttt{VINEWTONRSLS} solver in PETSc \cite{Balayetal2023}, denoted as \emph{RS Newton}.  We use it with a fixed number of preconditioned Krylov iterations to solve the arising linear systems.  For the symmetric Laplacian in Example \ref{ex:results:classical} we apply conjugate gradients (CG) with incomplete Cholesky (IC) preconditioning, but for other operators we apply the generalized minimum residual (GMRES) method with incomplete LU (ILU) preconditioning.  This requires $O(m_j)$ work per iteration. The coarse solver is also RS Newton, but iterated to convergence and with direct solvers applied to solve the arising linear systems.

@@

\medskip\noindent
Ed: \cite{GraeserKornhuber2009} only considers the Laplacian lower obstacle case, and they write everything in terms of nonlinear Gauss-Seidel (i.e.~optimize one basis function at a time).  They first lay out the \cite{Tai2003} CD strategy as Alg. 4.7, only for V(1,0) cycles.  Then they observe that the convergence rates from \cite{Tai2003} are slower (worse) than if one does "monotone multigrid".  This seems to mean to them that one modifies the coarse basis functions according to whether they overlap the current fine-grid estimate of the coincidence set, "truncated nodal basis functions".  (See the start of \cite{GraeserKornhuber2009} section 5.2.)  I think that making these modifications in Firedrake, so we could try monotone multigrid V-cycles, would be a nightmare.  FASCD does not attempt the monotone multigrid idea at all.
